 The Markdown File
=================================================================================================
 Loading and Preprocessing the data
 
```{r echo = TRUE}
require(lubridate)
activitydata <- read.csv("C:/Big Data and Hadoop/R/R Tutorial/course5 assignments1/activity.csv", stringsAsFactors = FALSE)
summary(activitydata)
activitydata$date<- ymd(activitydata$date)
```

 Mean total number of steps taken per day
 Calculate the total number of steps taken per day
 A histogram of the total number of steps taken each day
 Calculate and report the mean and median of the total number of steps taken per day
 

```{r echo=TRUE}
StepsbyDay<-tapply(activitydata$steps[!is.na(activitydata$steps)],
              activitydata$date[!is.na(activitydata$steps)],sum)

StepsbyDay

hist(StepsbyDay, main="Histogram of Steps Taken Each Day", xlab="Steps")

StepsbyDayMean<-mean(StepsbyDay[!is.na(StepsbyDay)])
StepsbyDayMedian<-median(StepsbyDay[!is.na(StepsbyDay)])

barplot(c(StepsbyDayMean,StepsbyDayMedian), 
        main="Mean/Median of Daily Steps Taken",col = c("red","green"),
        names.arg=c("Mean","Median"),ylim=c(10760,10770))
```

The average daily activity pattern
Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
dailyactptrn<-tapply(activitydata$steps[!is.na(activitydata$steps)],
                   activitydata$interval[!is.na(activitydata$steps)],mean)

par(mfrow=c(1,1))
plot(dailyactptrn, main="Time series of Steps Taken By Interval",type = "l",col="red")

maxInterval=data.frame(dailyactptrn[dailyactptrn==max(dailyactptrn)])
maxInterval

```

Imputing missing values
There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
A new dataset that is equal to the original dataset but with the missing data filled in.
Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r echo=TRUE}
intervalNames<-data.frame(rownames(data.frame(dailyactptrn)))
colnames(intervalNames)<-"interval"

IntervalAverages<-data.frame(c(data.frame(rownames(data.frame(dailyactptrn))),
                               data.frame(dailyactptrn)))
colnames(IntervalAverages)<-c("interval","Average_Interval_Steps")

RowsAll=length(activitydata$steps)
RowsNA=length(activitydata$steps[is.na(activitydata$steps)])

par(mfrow=c(1,1))
barplot(c(RowsNA,RowsAll), main="Number of Missing Rows",col = c("blue","grey"))

ActivityDataImputed <- merge(activitydata,IntervalAverages,by="interval")

ActivityDataImputed[is.na(ActivityDataImputed$steps),]$steps <-ActivityDataImputed[is.na(ActivityDataImputed$steps),]$Average_Interval_Steps

byDayImputed<-tapply(ActivityDataImputed$steps,ActivityDataImputed$date,sum)
hist(byDayImputed, main="Steps Taken Each Day With Missing as Avg for each Interval")

byDayMeanImputed<-mean(byDayImputed[!is.na(StepsbyDay)])
byDayMedianImputed<-median(byDayImputed[!is.na(StepsbyDay)])

par(mfrow=c(1,1))
barplot(c(byDayMeanImputed,byDayMedianImputed), 
        main="Mean/Median of Daily Steps and Missing Values as Avg for each Interval",col = c("blue", "green"))

ActivityDataImputed$weekdayInd<-with(ActivityDataImputed, !(weekdays(as.Date(ActivityDataImputed$date))=="Saturday" | weekdays(as.Date(ActivityDataImputed$date))=="Sunday")) 

```

Are there differences in activity patterns between weekdays and weekends
A new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r echo=TRUE}
ActivityDataImputed$TimePeriod[ActivityDataImputed$weekdayInd==0]<-"Weekend"
ActivityDataImputed$TimePeriod[ActivityDataImputed$weekdayInd==1]<-"Weekday"

library(dplyr)

ActivityDataImputed.grpdata <- group_by(ActivityDataImputed, TimePeriod,interval)
ActivityDataImputed.grpdata<-summarize(ActivityDataImputed.grpdata, AvgSteps = mean(steps, na.rm = TRUE))

library(lattice)
xyplot(AvgSteps ~ interval | TimePeriod, data = ActivityDataImputed.grpdata, 
        layout = c(1, 2),type = "l")

```

